Robust Clustering for Ad Hoc Cognitive Radio Network

### 2017-03-26 journal-draft/bare_jrnl_comsoc.tex
minor changes
first 5 pages modified

In CRN, clusters are formed in the very beginning of the network operation, and re-formed periodically according to the dynamics of the CRN.
Each formed cluster has one or multiple unlicensed channels which are available for every CR node in the cluster.
The available unlicensed channels are referred to in the following of this paper as \textit{common licensed channels} (or common channels for short, which is abbreviated as CC).
Both payload and control overheads can be transmitted on the CCs.
%Out of the available CCs, there are always one or multiple channels which are used for the payload communication.
When one or several cluster members can not access one certain CC on which primary user activity is detected, the channel will be excluded from the set of CCs.
In particular, if that channel is being used for payload communication, the communication pair will stop and resume the transmission on another available CC.
The availability of CCs within a cluster defines the existence of that cluster, \ie no CCs are available means the corresponding cluster doesn't exist.
%As long as there is a CC available, the corresponding cluster can continue to exchange payload data.
In the context of CRN, the activity of primary users is usually unknown to the secondary users, then the primary users' activity is deemed as random.
In this case, the cluster which secures more CCs will anticipate a longer life expectancy, in other words, is more robust.
It is obvious that fewer secondary users in one cluster yield more CCs, but this contradicts to one of the motivations of clustering, \ie the cooperative decision making, where more users result in more accurate sensing result~\cite{Consensus_based_clustering12}.
Thus the number of small clusters, especially the \textit{singleton clusters}, \ie the cluster which has only one CR node, should be minimized.
As to the cluster which consists of many nodes, although the spectrum sensing is benefited, it usually leads to less common channels, which undermines the robustness of the cluster.
Thus the cluster robustness discussed in terms of number of CCs carries little meaning when the sizes of formed clusters are not given consideration.
Cluster size also plays a part in transmission power consumption~\cite{clustering_globecom11, EnergyEfficientClusteringRouting_2015}.
In this paper we focus on the efficient formation of the clusters and the evaluation of the robustness of the formed clusters, but our scheme is also able to generate the clusters with the desired size, which gives the network designers more freedom when the power consumption is involved in the design of network applications.
%In this paper we will introduce a robust cluster structure for CRN, where the robustness means the ability of the clusters to sustain under the influence of the increasing activity of primary users, and the direct metric is the number of CCs in the clusters.

@@

There has been a lot of research done on clustering in wireless networks.
As there is no restriction on the nodes to access the spectrum in wireless ad-hoc, mesh networks and sensor networks, the objective of clustering is to decrease the transmission power consumption~\cite{Kawadia03}, to improve the routing performance~\cite{clustering_mesh_globecom2010}, or to imrove the network lifetime and coverage~\cite{Abbasi_survey_07}.
%Various clustering schemes are proposed to target different aspects in cognitive radio networks.
%Work~\cite{Consensus_based_clustering12} improves spectrum sensing ability by grouping the CR users with potentially best detection performance into the same cluster.
%Clustering scheme~\cite{clustering_globecom11} obtains the best cluster size which minimizes power consumption caused by communication within and among clusters.
%\cite{clustering_globecom11} proposes clustering strategy in cognitive radio network, which looks into the relationship between cluster size and power consumption and accordingly controlling the cluster size to decrease power consumption.
With regard to forming clusters in CRN, deciding on the common channel within each cluster is the foremost question to answer.
\cite{Zhao07, Chen07,Affinity_clustering_09icccn} propose the clustering schemes and make sure every cluster possesses at least one CC.
Clustering scheme~\cite{Consensus_based_clustering12} looks for the network partition which improves the accuracy of spectrum sensing with the cluster structure.
%Clustering scheme~\cite{clustering_globecom11} obtains the best cluster size which minimizes power consumption caused by communication within and among clusters.
\cite{TWC2012_cooperative_communication} forms the clusters by deciding on the cluster heads, where the transmit power for the long-haul transmission between the cluster heads is minimized.
\cite{clustering_globecom11} proposes a cluster structure which promises energy efficiency.
\cite{cluster_EW10} proposes strategy on how to decide on the CCs and access the multiple CCs within clusters.
An event-driven clustering scheme is proposed for cognitive radio sensor network in \cite{Ozger_cluster_crsn_13}.
No one among the above mentioned schemes provides the robustness to the formed clusters against primary users.

@@

We consider a set of CR users $\mathcal{N}$ and a set of primary users distributed over a given area.
A set of licensed channels $\mathcal{K}$ is available for the primary users. 
The CR users are allowed to transmit on channel $k \in \mathcal{K}$ only if no primary user is detected on channel $k$. 
CR users conduct spectrum sensing independently and sequentially on all licensed channels.\footnote{We assume that every node can detect the presence of an active primary user on each channel with certain accuracy. The spectrum availability can be validated with a certain probability of detection. Spectrum sensing/validation is out of the scope of this paper.}
We adopt the unit disk model~\cite{unitDiskModel} for both primary and CR users' transmission.
%As to the relation between the primary and secondary users,
If a CR node $i$ locates within the transmission range of an active primary user $p$, $i$ is not allowed to use the channel which is being used by $p$.
%On the other hand, if $i$ is not in the transmission range of primary user $p$, $i$ can certainly not detect the presence of $p$.
We assume the primary users change their operation channels slowly thus we omit the time index for the spectrum availability. 
As the result of spectrum sensing, $K_i \subseteq \mathcal{K}$ denotes the set of available licensed channels for CR user $i$.
As the transmission range of primary users is limited and secondary users have different locations, different secondary users have different views of the spectrum availability, i.e., for any $i, j \in \mathcal{N}$, $K_i = K_{j}$ does not necessarily hold.
We therefore represent the network of CR nodes by a graph $G = (\mathcal{N}, E)$, where $E \subseteq \mathcal{N} \times \mathcal{N}$ such that $\{i, j\} \in E$ if and only if $K_{i} \cap K_{j}\neq \emptyset$ and $d_{i,j} < r$, where $d_{i,j}$ is the distance between $i, j$ and $r$ is the radius of secondary user's transmission range. 
%
Among the secondary users, we denote by $\text{Nb}(i)$ user $i$'s neighborhood, which consists of the CR nodes located within the transmission range of $i$.

@@

We give the definition of cluster in CRN as follows. 
A cluster $C$ is a set of secondary nodes which possess the same set of common channels.
In particular, a cluster consists of a cluster head $h(C)$ and a number of cluster members, and the cluster head is able to communicate with any cluster member directly.
A cluster can be formed only by the cluster head.
%$\text{Nb}(i)$ denotes node $i$'s neighborhood which consists of all its one hop neighbors.
%For any cluster member $i \in C$, $i \in \text{Nb} (h_C) $ holds.
The size of $C$ is denoted by $|C|$.
When the cluster head of a cluster is $i$, we denote that cluster by $C(i)$.
$K(C)$ denotes the set of CCs in cluster $C$, $ K(C) = \Cap_{i\in C} K_i$.
%Clustering is performed periodically, because the secondary users are mobile and the primary users change their operation channels, thus the channel availability on secondary users changes accordingly.
The notations used in the system model are listed in Table~\ref{tab1}.
\begin{table}[h!]
\caption{Notations}
\label{tab1}
\centering
\begin{tabular}{llr}
\toprule
%\multicolumn{2}{c}{Item} \\
%\cmidrule(r){1-2}
Symbol & Description \\
\midrule
$\mathcal{N}$  & set of CR users in a CRN\\
$N$ & number of CR users in a CRN, $N=|\mathcal{N}|$\\
$\mathcal{K}$	& set of licensed channels\\
$k(i)$ & the working channel of user $i$\\
$\text{Nb}(i)$ & the neighborhood of CR node $i$    \\
$C(i)$ & a cluster whose cluster head is $i$  \\
%$k_i$   & set of available channels at CR node $i$  \\
$K_i$   & the set of available channels at CR node $i$  \\
$K(C(i))$   & the set of available CCs of cluster $C(i)$ \\
%$h(C)$ & cluster head of a cluster $C$\\
$h(C)$ & the cluster head of a cluster C\\
%$\text{CH}$ & cluster head\\
%$\text{CH}$ & cluster head\\
$\delta$ & the cluster size which is preferred\\
$S_i$ & a set of claiming clusters, each of which includes \\
& debatable node $i$ after phase I\\
$d_i$  & individual connectivity degree of CR node $i$\\
$g_i$  & neighborhood connectivity degree of CR node $i$\\
$f(C)$ & the number of CCs of a cluster $C$, which is used \\
& in the problem description\\
% $\mathcal{G}$ & a collection of some possible clusters in $\mathcal{N}$\\
 $\mathcal{S}$ & the collection of all the possible clusters in $\mathcal{N}$\\
 $C_i$  & the $i$th cluster in $\mathcal{S}$ \\
\bottomrule
\end{tabular}
\end{table}

@@

Given a cognitive radio network $\mathcal{N}$ where $|\mathcal{N}|=N$, the collection of all the possible clusters\footnote{Possible cluster means the collection of CR nodes, which complies with the definition in~\ref{sec:model}} is denoted as $\mathcal{S}$ where $\mathcal{S}=\{C_1, C_2,\ldots,C_{|\mathcal{S}|}\}$ \footnote{The subscripts of the clusters can be decided in any convenient way, \ie the sequence of identifying them.} and there is $\bigcup_{1\leq i \leq |\mathcal{S}|} C_i = N$.
%
When the requirements on the cluster size are enforced, \ie the desired size is $\delta$ and the cluster sizes should fall in the scope $\langle\delta_1, \delta_2\rangle$, where $\delta, \delta_1, \delta_2\in \mathbb{Z}^+$, $\delta_1 \leq \delta\leq \delta_2$, then a feasible clustering solution is a subcollection $\mathcal{S}' \subseteq \mathcal{S}$, which satisfies 
$\delta_1\leq|C_i|\leq \delta_2, \bigcup_{C_i\in \mathcal{S}'} C_i = \mathcal{N}$ and $C_{i'}\cap C_i =\emptyset$, $C_{i'}, C_i\in \mathcal{S}'$, $i'\neq i$.

The robust clustering problem in CRN is to find the optimal clustering solution from the feasible clustering solutions, whose sum of the numbers of CCs of clusters $\sum_{C\in \mathcal{S}'} f(C)$ is the maximal.

@@

\section{Centralized Solution for Robust Clustering}
\label{centralized_solution}

Based on Definition~\ref{def_centralized_clustering}, the decision version of this problem is to determine whether there is a non-empty $\mathcal{S}'\subseteq \mathcal{S}$, so that $\sum_{C\in \mathcal{S}'} f(C) \geqslant \lambda$ where $\lambda$ is a real number.
We have the following theorem on the complexity of this problem.

\begin{theorem}
\label{theorem1}
Robust clustering problem in CRN is NP-hard, when the maximum size of clusters is larger than 3, $\delta_1=1$ and $\delta_2 = N$.
%Assume a CRN can be represented by a connected graph, and there is at least one common channel between any pair of neighbours, then forming at least two CR nodes into one cluster is NP-complete.
\end{theorem}
The proof is in Appendix~\ref{proof_theorem1}.

We propose a centralized solution for the robust clustering problem, which solves an optimization with standard solvers. 
To formulate the optimization, we need to do some preparation beforehand.
First, all the possible clusters complying with the description in the system model are found and constitute a set $\mathcal{S}$.
Second, we assign a cluster size related weight to each cluster, which is positively correlated with the difference between the cluster size and the desired cluster size. 
Considering $|\mathcal{S}|=M$, $C_i\in \mathcal{S}$ means the $i$ th cluster in $\mathcal{S}$, and $\delta$ is the desired size, the size related weight for each cluster is given as follows,
$$
p_i(C_i) = \left\{ \begin{array}{rl}
0 &\mbox{ if $|C_i|=\delta$} \\
\rho_1 &\mbox{if $||C_i|-1|=\delta$} \\
\rho_2 &\mbox{if $||C_i|-2|=\delta$} \\
\vdots
\end{array} \right.
$$
where $\rho_1, \rho_2, \cdots$ are positive values.
In particular, $\rho$ increases with the divergence between $|C_i|$ and $\delta$, \ie these is $\rho_2> \rho_1>0$.

The optimization searches the set $\mathcal{S}$ and decides on certain clusters which constitute the whole CRN without overlapping between any two of them, besides, the sum of CCs of the chosen clusters is maximized.
Then, a central node (or controller) with the knowledge of all CR nodes (and also the possible clusters) can solve the problem based on the following formulation.

@@

\begin{equation}
\begin{aligned}
     &\max\limits_{w_i, x_{ij}} && \Sigma_{j=1}^N\Sigma_{i=1}^M (w_i\cdot t_{ij}) \\
     &\text{subject to}   && \Sigma_{i=1}^M x_{ij} = 1, for\,\,\forall j=1, \ldots, N \\
   &&& \Sigma_{j=1}^N x_{ij} = |C_i|\cdot w_i,\, for\,\,\forall i=1, \ldots, M \\
  % &&& \text{$x_{ij}$ and $w_i$ are binary variables.}\\
   &&& i\in \{1,2, \cdots M\}, \hspace{0.3cm} j\in \{1,2,\cdots N\}
%\notag
\end{aligned}
\label{centralized_opt}
\end{equation}

%The optimization is a binary linear programming problem.
The objective is to maximize the sum of CCs of all the clusters.
%We minimize the opposite of the real objective to make this problem solvable.
$w_i$ and $x_{ij}$ are the two binary variables in this problem.
% where $j\in \{1,2,\cdots,N-1, N\}$ and $i\in \{1,2,\cdots,M-1, M\}$.
$N$ is the total number of CR users in network $\mathcal{N}$, $M$ is the number of possible clusters in $\mathcal{S}$.
%$G = |\mathcal{G}|$ and there is $\mathcal{G} \subseteq \mathcal{S}$.
Being either 1 or 0, $w_i$ denotes whether the $i$th cluster $C_i$ in $\mathcal{S}$ is chosen by the solution or not.
$x_{ij}$ indicates whether the CR node $j$ resides in the $i$th potential cluster, \ie $x_{ij}=1$ means node $j$ resides in the cluster $C_i$. 
Node index $j$ is identical to the node ID.
%In practice, we can solve the problem with $\mathcal{G} = \mathcal{S}$ when $|\mathcal{S}|$ is not large. 
%
%
%Given a CRN $\mathcal{N}$ and desired cluster size $\delta$, we obtain a collection of clusters $\mathcal{G}$ which contains all the \textit{potential} clusters, and the sizes of these clusters are $1,2,\ldots,\delta$.
%Potential clusters are the clusters which satisfy the conditions introduced in Section~\ref{sec:model}. 
%Note that the potential clusters include the singleton ones.
%By permitting the existence of the singleton clusters, we can ensure that the $\mathcal{S}'$ in Definition~\ref{def_centralized_clustering} is always be feasible.
$t_{ij}$ is a constant which is $\frac{q_{ij}}{|C_i|} - p_i(C_i)$.
$q_{ij}= |K(C_i)|$ when there is $j\in C_i$, and $q_{ij}= 0$ when there is $j\notin C_i$.
$|K(C_i)|$ is the number of CCs of cluster $C_i$, and $|C_i|$ is the size of cluster $C_i$.
$p_i(C_i)$ is the size related weight which has been introduced.
%%Based on the knowledge on $\mathcal{S}$, we construct a $M\times N$ matrix $Q_{M\text{x}N}$ which is shown in Figure~\ref{costant_matrix_Q}. 
%%Constant $q_{ij}$ is the element of $Q_{M\text{x}N}$, and its subscripts correspond to the $i$th cluster and CR node $j$ respectively.
%$q_{ij}= |K(C_i)|/|C_i|$ when there is $j\in C_i$, and $q_{ij}= 0$ when there is $j\notin C_i$.
%In other words, each non-zero element $q_{ij}$ means the number of CCs of the cluster $i$ where node $j$ resides.

@@

Now we examine the objective function to see whether it in line with the goal to maximize the total number of CCs meanwhile consider the restriction on the cluster size.
Replacing $t_{ij}$ with $\frac{q_{ij}}{|C_i|} - p_i(C_i)$, the objective function becomes,
\begin{equation}
\begin{aligned}
     &\max\limits_{w_i, x_{ij}} && \Sigma_{j=1}^N\Sigma_{i=1}^M (w_i\cdot \frac{q_{ij}}{|C_i|} - w_i\cdot p_i(C_i)) \\
\notag
\end{aligned}
\label{centralized_opt}
\end{equation}
The sum of the first items is the sum of CCs of all the chosen clusters.
As to the second item, when $w_i$ is 1 ($C_i$ is chosen) and $|C_i|\neq\delta$, it will be negative, which contradicts the direction of the optimization.
Thus the second item discourages the appearance the clusters whose sizes are different from $\delta$, especially those whose sizes are not $\delta$.

@@

The constraints guarantee to obtain the clusters which together include all the CR users and don't overlap.
The first constraint regulates that each CR node should reside in exactly one cluster.
The second constraint regulates that when the $i$th possible cluster $C_i$ is chosen, there will be exactly $|C_i|$ CR nodes residing in $C_i$.

@@

This problem is a binary linear programming problem, which can be solved by many available solvers.
The difficulty of using this method lies in obtaining the set $\mathcal{S}$.
In the worst case, \ie the CRN forms a full connected graph, the size of $S$ is $\Sigma_{r=1}^{N}\ {N \choose r} = 2^N-1$.
To levitate this problem, a smaller set, \ie $\mathcal{G} \subset \mathcal{S}$ can be used.
In this case, the global optimal may not be found.
$\mathcal{G}$ can be prepared based on the cluster size, and it is recommended to include all the singleton clusters to make sure the availability of feasible solutions.

Another obstacle to apply this centralized scheme is, the centralized entity firstly needs to collect the information from all the CR nodes, then computes the clustering solution and distributes it across the whole network.
This process involves a large number of communication overheads.
In CRN, it is necessary to do clustering again on some occasions.
For example, when a certain amount of CR users move away from their clusters, \ie they loose the direction connection with any member in their previous clusters, or a certain amount of clusters can not be maintained as the CCs in these clusters don't exist any longer due to primary users' activity. 
Hence, when the spectrum availability and the CR users' location change frequently, the centralized robust clustering scheme is not suitable for CRN.

>>>>

In CRN, clusters are formed in the very beginning of the network operation, and re-formed periodically according to the dynamics of the CRN.
Each formed cluster has one or multiple unlicensed channels which are available for every CR node in the cluster.
The available unlicensed channels are referred to in the following of this paper as \textit{common licensed channels} (or common channels for short, which is abbreviated as CC).
Both payload and control overheads can be transmitted on the CCs.
%Out of the available CCs, there are always one or multiple channels which are used for the payload communication.
When one or several cluster members can not access one certain CC on which primary user activity is detected, the channel will be excluded from the set of CCs.
In particular, if that channel is being used for payload communication, the communication pair will stop and resume the transmission on another available CC.
The availability of CCs within a cluster defines the existence of that cluster, \ie no CCs are available means the corresponding cluster doesn't exist.
%As long as there is a CC available, the corresponding cluster can continue to exchange payload data.
In the context of CRN, the activity of primary users is usually unknown to the secondary users, then the primary users' activity is deemed as random.
In this case, the cluster which secures more CCs will anticipate a longer life expectancy, in other words, is more robust.
It is obvious that fewer secondary users in one cluster yield more CCs, but this contradicts to one of the motivations of clustering, \ie the cooperative decision making, where more users result in more accurate sensing result~\cite{Consensus_based_clustering12}.
Thus the number of small clusters, especially the \textit{singleton clusters}, \ie the cluster which has only one CR node, should be minimized.
As to the cluster which consists of many nodes, although the spectrum sensing is benefited, it usually leads to less common channels, which undermines the robustness of the cluster.
Thus the cluster robustness discussed in terms of number of CCs carries little meaning when the sizes of formed clusters are not given consideration.
Cluster size also plays a part in transmission power consumption, \ie cluster size affects the transmission power consumption when routing is conducted~\cite{clustering_globecom11, EnergyEfficientClusteringRouting_2015}.
In this paper we focus on the efficient formation of the clusters and the evaluation of the robustness of the formed clusters, but our scheme is also able to generate the clusters with the desired size, which gives the network designers more freedom when the power consumption is involved in the design of network applications.
%In this paper we will introduce a robust cluster structure for CRN, where the robustness means the ability of the clusters to sustain under the influence of the increasing activity of primary users, and the direct metric is the number of CCs in the clusters.

@@

Different clustering strategies have been proposed for wireless networks of different type.
As there is no restriction on the nodes to access the spectrum in wireless ad-hoc, mesh networks and sensor networks, the objective of clustering is to decrease the transmission power consumption~\cite{Kawadia03}, to improve the routing performance~\cite{clustering_mesh_globecom2010}, or to imrove the network lifetime and coverage~\cite{Abbasi_survey_07}.
%Various clustering schemes are proposed to target different aspects in cognitive radio networks.
%Work~\cite{Consensus_based_clustering12} improves spectrum sensing ability by grouping the CR users with potentially best detection performance into the same cluster.
%Clustering scheme~\cite{clustering_globecom11} obtains the best cluster size which minimizes power consumption caused by communication within and among clusters.
%\cite{clustering_globecom11} proposes clustering strategy in cognitive radio network, which looks into the relationship between cluster size and power consumption and accordingly controlling the cluster size to decrease power consumption.
With regard to forming clusters in CRN, deciding on the common channel within each cluster is the foremost question to answer.
\cite{Zhao07, Chen07,Affinity_clustering_09icccn} propose the clustering schemes and make sure every cluster possesses at least one CC.
Clustering scheme~\cite{Consensus_based_clustering12} looks for the network partition which improves the accuracy of spectrum sensing with the cluster structure.
%Clustering scheme~\cite{clustering_globecom11} obtains the best cluster size which minimizes power consumption caused by communication within and among clusters.
\cite{TWC2012_cooperative_communication} forms the clusters by deciding on the cluster heads, where the transmit power for the long-haul transmission between the cluster heads is minimized.
\cite{clustering_globecom11} proposes a cluster structure which promises energy efficiency.
\cite{cluster_EW10} proposes strategy on how to decide on the CCs and access the multiple CCs within clusters.
An event-driven clustering scheme is proposed for cognitive radio sensor network in \cite{Ozger_cluster_crsn_13}.
No one among the above mentioned schemes provides the robustness to the formed clusters against primary users.

@@

We consider a set of CR users $\mathcal{N}$ and a set of primary users distributed over a given area.
A set of licensed channels $\mathcal{K}$ is available for the primary users. 
The CR users are allowed to transmit on channel $k \in \mathcal{K}$ only if no primary user is detected on channel $k$. 
CR users conduct spectrum sensing independently and sequentially on all licensed channels.\footnote{We assume that every node can detect the presence of an active primary user on each channel with certain accuracy. The spectrum availability can be validated with a certain probability of detection. Spectrum sensing/validation is out of the scope of this paper.}
We adopt the unit disk model~\cite{unitDiskModel} for both primary and CR users' transmission.
%As to the relation between the primary and secondary users,
If a CR node $i$ locates within the transmission range of an active primary user $p$, $i$ is not allowed to use the channel which is being used by $p$.
%On the other hand, if $i$ is not in the transmission range of primary user $p$, $i$ can certainly not detect the presence of $p$.
We assume the primary users change their operation channels slowly, thus we consider the clustering problem at some point in time given some degree of information of the availability of the control channels.
Due to the same reason, we omit the time index for the spectrum availability. 
As the result of spectrum sensing, $K_i \subseteq \mathcal{K}$ denotes the set of available licensed channels for CR user $i$.
As the transmission range of primary users is limited and secondary users have different locations, different secondary users have different views of the spectrum availability, i.e., for any $i, j \in \mathcal{N}$, $K_i = K_{j}$ does not necessarily hold.
We therefore represent the network of CR nodes by a graph $G = (\mathcal{N}, E)$, where $E \subseteq \mathcal{N} \times \mathcal{N}$ such that $\{i, j\} \in E$ if and only if $K_{i} \cap K_{j}\neq \emptyset$ and $d_{i,j} < r$, where $d_{i,j}$ is the distance between $i, j$ and $r$ is the radius of secondary user's transmission range. 
%
Among the secondary users, we denote by $\text{Nb}(i)$ user $i$'s neighborhood, which consists of the CR nodes located within the transmission range of $i$.

@@

We give the definition of cluster in CRN as follows. 
A cluster $C$ is a set of secondary nodes which possess the same set of common channels.
In particular, a cluster consists of a cluster head $h(C)$ and a number of cluster members, and the cluster head is able to communicate with any cluster member directly.
A cluster can be formed only by the cluster head.
%$\text{Nb}(i)$ denotes node $i$'s neighborhood which consists of all its one hop neighbors.
%For any cluster member $i \in C$, $i \in \text{Nb} (h_C) $ holds.
The size of $C$ is denoted by $|C|$.
When the cluster head of a cluster is $i$, we denote that cluster by $C(i)$.
$K(C)$ denotes the set of CCs in cluster $C$, $ K(C) = \Cap_{i\in C} K_i$.
%Clustering is performed periodically, because the secondary users are mobile and the primary users change their operation channels, thus the channel availability on secondary users changes accordingly.
The notations used in the system model are listed in Table~\ref{tab1}.
\begin{table}[h!]
\caption{Notations}
\label{tab1}
\centering
\begin{tabular}{llr}
\toprule
%\multicolumn{2}{c}{Item} \\
%\cmidrule(r){1-2}
Symbol & Description \\
\midrule
$\mathcal{N}$  & set of CR users in a CRN\\
$N$ & number of CR users in a CRN, $N=|\mathcal{N}|$\\
$\mathcal{K}$	& set of licensed channels\\
$k(i)$ & the working channel of user $i$\\
$\text{Nb}(i)$ & the neighborhood of CR node $i$    \\
$C(i)$ & a cluster whose cluster head is $i$  \\
%$k_i$   & set of available channels at CR node $i$  \\
$K_i$   & the set of available channels at CR node $i$  \\
$K(C(i))$   & the set of available CCs of cluster $C(i)$ \\
%$h(C)$ & cluster head of a cluster $C$\\
$h(C)$ & the cluster head of a cluster C\\
%$\text{CH}$ & cluster head\\
%$\text{CH}$ & cluster head\\
$\delta$ & the cluster size which is preferred\\
$S_i$ & a set of claiming clusters, each of which includes \\
& debatable node $i$ after phase I\\
$d_i$  & individual connectivity degree of CR node $i$\\
$g_i$  & neighborhood connectivity degree of CR node $i$\\
$f(C)$ & the number of CCs of a cluster $C$, which is used \\
& in the problem description\\
% $\mathcal{G}$ & a collection of some possible clusters in $\mathcal{N}$\\
 $\mathcal{S}$ & the collection of all the possible clusters in $\mathcal{N}$\\
 $C_i$  & the $i$-th cluster in $\mathcal{S}$ \\
 $|C_i|$ & size of the cluster $C_i$\\
 $|K(C_i)|$ & the number of CCs of cluster $C_i$\\
\bottomrule
\end{tabular}
\end{table}

@@

As to a cognitive radio network, where the set of CR nodes is $\mathcal{N}$.
Each cluster complies with the definition in~\ref{sec:model}, whose number of CCs in a cluster $C$ is $f(C)$.
The robust clustering problem is to decide the set of clusters $\mathcal{T}$, where 
\begin{enumerate}
\item the sum of the $f(C) is the maximal$, where $C\in \mathcal{T}$ 
\item the intersection of any two clusters in $\mathcal{T}$ is an empty set
\item the union of clusters in $\mathcal{T}$ is $\mathcal{N}$
\item the cluster sizes should fall in the scope $\langle\delta_1, \delta_2\rangle$, where $\delta, \delta_1, \delta_2\in \mathbb{Z}^+$, $\delta_1 \leq \delta\leq \delta_2$, and $\delta$ is decided according to the application to be implemented.
\end{enumerate}

Assume we have obtained the set of $\mathcal{S}$ which contains all the possible clusters in $\mathcal{N}$, \ie $\mathcal{S}=\{C_1, C_2,\ldots,C_i, \ldots, C_{|\mathcal{S}|}\}$ \footnote{The subscript $i$ means the $i$-th cluster in $\mathcal{S}$.} and there is $\bigcup_{1\leq i \leq |\mathcal{S}|} C_i = \mathcal{N}$.
The decision version of this problem is to determine whether there is a non-empty set $\mathcal{S}'\subseteq \mathcal{S}$, so that $\sum_{C_i\in \mathcal{S}'} f(C_i) \geqslant \lambda$ where $\lambda$ is a real number.
We have the following theorem on the complexity of this problem.

\begin{theorem}
\label{theorem1}
The robust clustering problem in CRN is NP-hard, when the maximum size of clusters is larger than 3, $\delta_1=1$ and $\delta_2 = N$.
%Assume a CRN can be represented by a connected graph, and there is at least one common channel between any pair of neighbours, then forming at least two CR nodes into one cluster is NP-complete.
\end{theorem}
The proof is in Appendix~\ref{proof_theorem1}.

@@

\section{Centralized Solution for Robust Clustering}
\label{centralized_solution}
When the global knowledge of the CRN is available to us, we can propose a centralized scheme as comparison.
The proposed centralized solution formulates the problem in Definition~\ref{def_centralized_clustering} as an optimization problem, then the latter is solved with standard software packages.
The optimization searches the set $\mathcal{S}$ and decides on the clusters according to the following optimization formulation,
%
%which constitute the whole CRN without overlapping between any two of them, besides, the sum of CCs of the chosen clusters is maximized.
%Then, a central node (or controller) with the knowledge of all CR nodes (and also the possible clusters) can solve the problem based on the following formulation. 
%
%Second, we assign a cluster size related weight to each cluster, which is positively correlated with the difference between the cluster size and the desired cluster size.

@@

\begin{equation}
\begin{aligned}
     &\max\limits_{w_i, x_{ij}} && \Sigma_{j=1}^N\Sigma_{i=1}^M (w_i\cdot t_{ij}) \\
     &\text{subject to}   && \Sigma_{i=1}^M x_{ij} = 1, for\,\,\forall j=1, \ldots, N \\
   &&& \Sigma_{j=1}^N x_{ij} = |C_i|\cdot w_i,\, for\,\,\forall i=1, \ldots, M \\
  % &&& \text{$x_{ij}$ and $w_i$ are binary variables.}\\
   &&& i\in \{1,2, \cdots M\}, \hspace{0.3cm} j\in \{1,2,\cdots N\}
%\notag
\end{aligned}
\label{centralized_opt}
\end{equation}
This problem is a binary linear programming problem, which can be solved by many available solvers.
$w_i$ and $x_{ij}$ are two binary variables.
% where $j\in \{1,2,\cdots,N-1, N\}$ and $i\in \{1,2,\cdots,M-1, M\}$.
%$G = |\mathcal{G}|$ and there is $\mathcal{G} \subseteq \mathcal{S}$.
Being either 1 or 0, $w_i$ denotes whether the $i$-th cluster $C_i$ in $\mathcal{S}$ is chosen or not.
$x_{ij}$ indicates whether the CR node $j$ resides in the cluster $C_i$, \ie $x_{ij}=1$ means node $j$ resides in the cluster $C_i$. 
%Node index $j$ is identical to the node ID.
$N$ is the total number of CR users in network $\mathcal{N}$, $M$ is the number of possible clusters in $\mathcal{S}$.

The constraints guarantee to obtain the clusters which together include all the CR users and don't overlap.
The first constraint regulates that a CR node should reside in exactly one cluster.
The second constraint regulates that when the $i$-th cluster $C_i$ is chosen, there will be exactly $|C_i|$ CR nodes residing in $C_i$.

%The optimization is a binary linear programming problem.
The objective is to maximize the sum of the numbers of CCs in the clusters which constitute the whole CRN.
$t_{ij}$ is a constant and there is,
%In practice, we can solve the problem with $\mathcal{G} = \mathcal{S}$ when $|\mathcal{S}|$ is not large. 
%
%
%Given a CRN $\mathcal{N}$ and desired cluster size $\delta$, we obtain a collection of clusters $\mathcal{G}$ which contains all the \textit{potential} clusters, and the sizes of these clusters are $1,2,\ldots,\delta$.
%Potential clusters are the clusters which satisfy the conditions introduced in Section~\ref{sec:model}. 
%Note that the potential clusters include the singleton ones.
%By permitting the existence of the singleton clusters, we can ensure that the $\mathcal{S}'$ in Definition~\ref{def_centralized_clustering} is always be feasible.

\begin{equation}
t_{ij} = \frac{q_{ij}}{|C_i|} - p_i(C_i)
\end{equation}
where $q_{ij}= |K(C_i)|$ when node $j\in C_i$, and $q_{ij}= 0$ when node $j\notin C_i$.
%$|K(C_i)|$ is the number of CCs of cluster $C_i$, and $|C_i|$ is the size of cluster $C_i$.
$p_i(C_i)$ is the size-related weight, which reflects the deviation between the size of $C_i$ and the desired size.
Assuming $\delta$ is the desired size, then
$$
p_i(C_i) = \left\{ \begin{array}{rl}
0 &\mbox{ if $|C_i|=\delta$} \\
\rho_1 &\mbox{if $||C_i|-1|=\delta$} \\
\rho_2 &\mbox{if $||C_i|-2|=\delta$} \\
\vdots
\end{array} \right.
$$
where $\rho_1, \rho_2, \cdots$ are positive values and these is $\rho_2> \rho_1>0$.

%%Based on the knowledge on $\mathcal{S}$, we construct a $M\times N$ matrix $Q_{M\text{x}N}$ which is shown in Figure~\ref{costant_matrix_Q}. 
%%Constant $q_{ij}$ is the element of $Q_{M\text{x}N}$, and its subscripts correspond to the $i$th cluster and CR node $j$ respectively.
%$q_{ij}= |K(C_i)|/|C_i|$ when there is $j\in C_i$, and $q_{ij}= 0$ when there is $j\notin C_i$.
%In other words, each non-zero element $q_{ij}$ means the number of CCs of the cluster $i$ where node $j$ resides.

@@

Now we examine the objective function to see whether it in line with the goal to maximize the total number of CCs meanwhile consider the restriction on the cluster size.
Replacing $t_{ij}$ with $\frac{q_{ij}}{|C_i|} - p_i(C_i)$, the objective function becomes,
\begin{equation}
\begin{aligned}
     &\max\limits_{w_i, x_{ij}} && \Sigma_{j=1}^N\Sigma_{i=1}^M (w_i\cdot \frac{q_{ij}}{|C_i|} - w_i\cdot p_i(C_i)) \\
\notag
\end{aligned}
\label{centralized_opt}
\end{equation}
The sum of the first items is the sum of CCs of all the chosen clusters.
As to the second item, when $w_i$ is 1 ($C_i$ is chosen) and $|C_i|\neq\delta$, it will be negative, which contradicts the direction of the optimization.
Thus the second item discourages the appearance the clusters whose sizes deviate from $\delta$.

@@

@@

The difficulty of using this method lies in obtaining the set $\mathcal{S}$.
In the worst case, \ie the CRN forms a full connected graph, the size of $\mathcal{S}$ is $\Sigma_{r=1}^{N}\ {N \choose r} = 2^N-1$.
%To levitate this problem, a smaller set, \ie $\mathcal{G} \subset \mathcal{S}$ can be used.
%In this case, the global optimal may not be found.
%$\mathcal{G}$ can be prepared based on the cluster size, and it is recommended to include all the singleton clusters to make sure the availability of feasible solutions.

Another obstacle to apply this centralized scheme comes from the characteristics of CRN.
In order to apply this scheme in practice, a large over head needs to be factored in in collecting the channel usage as seen by the R nodes, as well as the connectivity between the CR nodes.
After computing the clustering solution, the centralized node needs to distribute it across the whole network.

In CRN, it is necessary to do clustering again on some occasions.
For example, when a certain amount of CR users move away from their clusters, \ie they loose the direction connection with any member in their previous clusters, or a certain amount of clusters can not be maintained as the CCs in these clusters don't exist any longer due to primary users' activity. 
Hence, when the spectrum availability and the CR users' location change frequently, the centralized robust clustering scheme is not suitable for CRN.
